Anxiety: a decision-theoretic perspective

### 2019-02-03 manuscript/manuscript.tex
making progress, making progress

\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript, palatino}
\usepackage{graphicx}
\usepackage{amsfonts, amsmath}
\usepackage{algorithm, algpseudocode}%

@@

\subsection{Robust Control}
In the previous section, we observed that the value of a state and its associated
state-action pairs are a function of the environment transition distribution and
the best future action. In standard MDP problems, it is typically assumed that the
transition distribution is known and accurate, and that the agent is consistently
able to operate under the optimal policy in the future. In this section, we introduce
a formalism deviating from these standard assumptions.

@@

In some instances, we solve for Q-values and state values directly through value
iteration. This is a form of Dynamic Programming. In other instances, we estimate
these terms through temporal difference learning. We use beta-pessimism (betamax)
learning for both. We use the former to solve for final (infinite horizon) values
and the latter to identify finite horizon values. We explain when we use both.

@@

One symptom central to all of the anxiety disorders is pessimistic inference,
or a tendency to appraise the likelihood and/or severity of threatening situations
out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
the probability of encountering their fear (e.g. a venomous spider) in their
everyday lives. The first studies of pessimistic inference in clinically anxious
populations involved individuals judging the perceived likelihood and severity
of different aversive situations, with the general finding that anxious individuals
rated these as more likely than non-clinical individuals \citep{ButlerMathews1983,
ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997,
Borkovec1999}; a finding that has been replicated more recently \citep{Maner2006,
Mitte2007, Miranda2007}.

In self-report,
individuals with anxiety rate hypothetical future negative outcomes as
being more likely to occur and worse to experience than non-anxious counterparts
\citep{ButlerMathews1983, ButlerMathews1987, Foa1996, MacLeod1996, MacLeod1997,
Luten1997, Stober1997, Borkovec1999, Maner2006, Mitte2007, Miranda2007}. In the
laboratory, individuals with anxiety demonstrate excessive fear in fear conditioning
paradigms. For example, anxiety is correlated with exaggerated fear towards
non-threatening cues during conditioning and extinguished cues during extinction
\citep{lissek2005, Duits2015}.

Experimentally, aberrations in threat appraisal have been repeatedly observed
in a litany of associative learning studies. Common to all of these studies,
participants learn that a particular stimulus (CS+) predicts the occurrence
of a negative outcome (e.g. shock, loud noise, aversive image). In simple conditioning,
there is some evidence to suggest that individuals with anxiety exhibit elevated
levels of response to threatening stimuli \citep{lissek2005, MinekaOehlberg2008,
Duits2015}. Importantly, conditioning studies also suggest impaired safety learning
in anxiety, wherein a stimulus never paired with an aversive outcome (CS-) is also
feared \citep{Duits2015}. Other notable findings include impaired blocking,
conditioned discrimination (AX+/BX-) (citations), and conditioned inhibition
(A+, AB-) (citations).

% I think when you come back to edit this, you need to highlight these findings
% more in the context of what you are about to explain. Pessimistic inference
% is both making bad things worse and, most importantly, that non-bad things
% become bad. Although one question is how much to include the other less
% studied paradigms

@@

In our first result, we demonstrate one simple mechanism by which pessimistic
inference and exaggerated threat appraisals may arise. In the *open field* environment,
an agent is placed in a large grid world entirely empty except for two states,
leading to positive and negative outcomes, respectively. Under complete control
(objective or perceived), an agent in this environment in under no threat despite
the presence of negative states. This is simply because, in a fully controllable
world, an agent need not ever interact with the aversive state. In fact, an agent
can decide to walk straight up to the aversive state and, insofar it does not
choose to take one step further, will never be in any actual danger.

>>>>

environment\documentclass[11pt]{article} % For LaTeX2e
\usepackage{manuscript, palatino}
\usepackage{graphicx}
\usepackage{amsfonts, amsmath}
\usepackage{algorithm, algpseudocode}%

@@

\subsection{Pessimistic Learning}

In the previous section, we observed that the value of a state and its associated
state-action pairs are a function of the environment transition distribution and
the best future action. In standard MDP problems, it is typically assumed that the
transition distribution is known and accurate, and that the agent is consistently
able to operate under the optimal policy in the future. In this section, we introduce
a formalism deviating from these standard assumptions.

@@

In the following sections, we present the results of simulations of pessimistic
learning in a variety of MDP environments. In all simulations, the state-action
values under the optimal policy, $Q_*(s,a)$, were solved for directly through
Q-value iteration \citep{SuttonBarto1998,SuttonBarto1998,bertsekas2005}. The
details of this algorithm are presented in the appendix. Model parameters are
reported for each environment in their corresponding sections. All simulations
were carried out using the scientific python distribution and are publicly available
at on Github (https://github.com/szorowi1/SecretFunTimes).

@@

@@

In this first section, we demonstrate how learning under an assumption of a lack
of future control naturally leads to pessimistic inference and second-order conditioning.
A symptom central to all of the anxiety disorders is pessimistic inference,
or a tendency to appraise the likelihood and/or severity of threatening situations
out of proportion to their actual environmental contingencies \citep{dsm5, BeckClark1997,
ClarkBeck2011}. For example, an individual with specific phobia may vastly exaggerate
the probability of encountering their fear (e.g. a venomous spider) in their
everyday lives. Anxious individuals will also tend to associate fear with cues
tangentially related to signals of threat. For example, the same individual may
begin to fear locations where they may encounter spider webs (e.g. basements).

Pessimistic inference in anxiety is well-documented. In studies of self-reported beliefs,
individuals with anxiety rate
hypothetical future negative outcomes as being more likely to occur and worse to
experience than non-anxious counterparts \citep{ButlerMathews1983, ButlerMathews1987,
Foa1996, MacLeod1996, MacLeod1997, Luten1997, Stober1997, Borkovec1999, Maner2006,
Mitte2007, Miranda2007}. In the laboratory, exaggerated threat appraisal in anxiety
has traditionally been measured through fear conditioning. The results of these
studies demonstrate that anxious individuals exhibit greater fear towards threatening
cues, non-threatening cues, and extinguished threatening cues as compared to non-anxious
counterparts \citep{lissek2005, MinekaOehlberg2008, Duits2015}. Importantly, threatening
cues can become associated with secondary neutral cues, allowing negative value to
spread citep{wessa2007}. In sum, anxiety is associated with inflated estimates of
threat, both likelihood and value, and these estimates can spread from the source
to distantly associated cues.

In our first simulation, we demonstrate how this may come about. In the open field
environment (Figure 1a), an agent is free to navigate a large grid world devoid
of value except for two states: one rewarding and one aversive red. In this environment,
the one-step transition structure is deterministic such that the environment
fully respects the agent's choice. As such, the agent is under no threat despite
the presence of a harmful state. Insofar that the agent never acts under a
masochistic policy and intentionally approaches harm, the agent need not ever
encounter the harmful state.

The long-run value estimates learned by an optimistic agent (i.e. an agent expecting
to be able to follow a reward-maximizing policy now and in the future) reflect this
(Figure 1b). For an optimistic agent, all states in the open field environment,
sans the harmful state itself, take on a positive value. This is unsurprising
insofar that, except for the harmful state, the agent will always encounter
a reward in the long-run no matter its starting position in the field.
